\section{General}
\subsection{Membership}
\index{Policy}
\begin{enumerate}
\item 	The Society supports the right to a leave of absence for its members. Whereas the Society recognizes that departments may have different expectations of graduate students, the terms of leaves of absences are determined by each department separately. (Med ID: 337)
\item  The Society supports differential ancillary fees for graduate students commensurate with the level of service provided and consistent with the overall level of budget reductions and academic units at Western. 
\item 	The Society supports reducing graduate students ancillary fees, including fees paid to support Intercollegiate Athletics.
\item 	Special status students, such as exchange and International graduate students, who wish to gain access services offered by the Society may do so if: (1) they pay the Society's and the CFS membership fees, and (2) pay all fees normally associated with the SOGS program(s) they wish to access. 
\item 	The Society supports international students at UWO and advocates for fair and regulated international students fees. The Society will lobby and pressure the University to freeze international student tuition fees as is the case at other universities in Ontario. 
\item 	The Society supports the right to appeal. In any academic appeal, a graduate student may use the support of law students, peers, faculty members and or family members. 
\end{enumerate}

\subsection{The Executive}
\index{Policy}
\begin{enumerate}
\item 	The President and other Executives of the Society shall be compensated for extra training and for work-related duties performed in addition to their contractual duties. 
\end{enumerate}


\subsection{Finance and Banking}
\index{Policy}
\begin{enumerate}
\item The Society's fiscal year ends April 30th. 
\item	The Society will carry forward any unused or surplus dollars from its budget to the following year. The carry forward will be approved by the Executives of the Society. 
\item	The Society will contribute top-up funds to the Joint Fund for Graduate Scholarship and Research program to match the top-up funding provided by the Faculty of Graduate Studies and Research Western.
\item	The Society endorses the view that graduate student funding should guarantee's cost of living, indexed to CPI.
\end{enumerate}

\subsection{Duties of the Executive }
\index{Policy}
\begin{enumerate}
\item 	The Speaker and Committee Members shall review all final reports and office matters regarding all outgoing executive and non-executive members. Upon approval of reports, paychecks will be issued to the parties involved.
\item  The Council authorizes SOGS Executive to make investment decisions on behalf of the Society. 
\end{enumerate}


\subsection{Duties of Non-Executive Officers }
\index{Policy}
\begin{enumerate}
\item 	SOGS non-executive officers shall be allocated space in the SOGS office for the purpose of performing their SOGS related duties.  
\end{enumerate}
 
 
\subsection{Committees of the Society }
\index{Policy}
\subsubsection{Creating and Disbanding}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 	SOGS non-executive officers shall be allocated space in the SOGS office for the purpose of performing their SOGS related duties.  
\end{longenum}



\subsubsection{The Academic Committee}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 		Whereas the society promotes graduate student research and research dissemination via the Grad Cast. In an effort to promote environmentally friendly operations, the Journal should be published in an online (web-based) format only. 
\item  	The Society encourages the presence of all executives and non-executives at the Western Research Forum (WRF).
\end{longenum}


\subsubsection{The Finance Committee}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 	The President of the Society will serve as the SOGS representatives with negotiations to the University regarding the ongoing working of the Ontario Students Opportunity Trust Fund (OSOTF) programme.   
\end{longenum}


\subsubsection{The Grad Club Committee}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 	The Society supports environmentally friendly and sustainable operations. Consistent with this, the Society supports the Grad Club's composting program initiative.    
\end{longenum}

\subsubsection{Orientation Committee}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 	Orientation Committee is responsible for creating events to welcome new graduate students.  
\end{longenum}

\subsubsection{The Policy Committee}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item 	The purpose of the Policy Committee is to review motions and make a policy manual. The Policy Committee is responsible for collecting all standing committees' meeting agendas and minutes for storage in SOGS office. The Policy Committee is also responsible for making policy recommendations and amendments to the Policy Document and to do SOGS Constitution, which should be submitted as motions to the Council for approval.  
\end{longenum}

\subsection{Personnel}
\index{Policy}
\begin{enumerate}
\item 	The Society opposes reductions in vacation time for current employees. 
\item 	The Society's staff will review, approve and implement the staff's Registered Retirement Savings Plan (RRSP) plan. \index{Personnel!RRSP}
\item 	Members of the Society shall act in accordance with the Personnel Policy. 
\item 	Council will eliminate the requirement for the President to conduct performance reviews of the Executive. 
\item 	The Society supports revisions to and adoption of the Personnel Policy Manual. 
\item 	The Personnel Policy shall note the roles of an Accounting Manager, an Office Manager and Grad Club Manager. 
\item 	The Society supports access to the SOGS Health Plan for all Grad Club staff/employees. 
\item 	The Society supports Workers Compensation coverage for its employees. The Society will make budgetary adjustments to match increasing cost of coverage. 
\end{enumerate}


\subsection{Finance}
\index{Policy}
\begin{enumerate}
\item The Speaker and Committee Members shall review all final reports and office matters regarding all outgoing executive and non-executive members. Upon approval of reports, paychecks will be issued to the parties involved.  
\item The Society prohibited from spending any unbudgeted money.  
\item Audited financial statements will be made available in September if (a) VP Finance resignation leads to a delay OR (b) an internal audit in August delays the process.  
\item The society is committed to paying its Executive wages commensurate with the wages paid to TAs. 
\item Executive members of the Society shall be compensated in proportion to TA wages, as follows: 1/2 TA for VP's and 1.5 TA for the President.  
\end{enumerate}


\subsubsection{Under section 2.12.2}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item The Long-term Planning and Priorities Committee shall oversee responsible and appropriate spending of the Capital Reserves from the Grad Club. 
\item The SOGS Policy Manual clearly defines pay for Grad Club employees.  
\item The Administrative Assistant has signing authority on all SOGS accounts.  
\end{longenum}

\subsubsection{Under section 2.12.8.4}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item	The Society will be included in the Grad Club's commercial property and general liability insurance policy, provided an agreement on the insurance broker and premium is agreed upon by both the Society and Grad Club.

\end{longenum}

\subsubsection{Under section 2.12.8.2}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item Students who surrender their original bus pass within three weeks of receiving a replacement bus pass are eligible to receive a refund of the replacement charge less a \$5.00 administration fee.  
\end{longenum}

\subsubsection{Under section 2.12.8.2}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item	The Society will fund a portion of the van program with the refund provided by the LTC, and the remainder be refunded to graduate students. (NO ID - Row 130 Filtered) \end{longenum} 

\subsection{Graduate Support Programs}
\begin{enumerate}
\item	Whereas the Society supports regulation of student fees and opposes tuition fee increases, the Society supports setting student fees based on 12-month periods. \end{enumerate}

\subsection{Elections Policy }
\begin{enumerate}
\item		Concerns for or complaints of Council members engaging in unprofessional conduct shall be brought before a judiciary committee, the Elections Appeal Committee that will evaluate the allegations. 
\end{enumerate}

\subsection{Procurement}
\subsubsection{General}
\begin{longenum}[ label*=\thesubsection.\arabic*., align=left]
\item	The Society supports the use of generic business cards for its Executive members provided they are purchased at the most economical value. 
\item	The Society supports the University's Code of Conduct for Trademark Licensees doing business with the University of Western Ontario Bookstore.     
\end{longenum}

\subsubsection{Under section 3.5.17.4}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item  	The SOGS Daily Planner/Handbook Committee will work with Grad Club to ensure the Grad Club has priority for advertising space on the outside back cover.
\end{longenum}

\subsubsection{Under section 3.5.16.5}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item   	The Society supports the use of external legal professionals to resolve contracts issues and impeachment issues. Where SOGS is involved in a contract, all matters shall be handled under the advisement of a lawyer/legal professional. 
\item Where the Society is involved in a contract, a minimum of 6 weeks of debate and 2 council meetings are required before ratification. 
\end{longenum}

\subsubsection{Under section 3.5.15}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item   The Society will solicit competitive bids whenever purchases, agreements, or contracts are considered.  
\end{longenum}

\section{Operations Policies}

\subsection{Council Policies}
\subsubsection{3.6.1	Pertaining to Council Meetings}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item  	The society is committed to equality across all departments, and council seats will be determined by SGPS-generated codes. Departments from separate codes are free to form one graduate student association, provided they provide the request in writing.
\item 	Mailouts must follow the agenda.  
\item 	The chair of meetings has control, not the speaker. 
\item 	All main motions and amendments must be in writing.  
\item 	Councillors will be compensated for childcare 
\item 	Audio recordings will be taken to aid the Secretary in verifying minutes, and these recordings will not be available outside the SOGS office.  
\end{longenum}

\subsubsection{ 3.6.2. The UWO Constitution and Bylaws}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item   	Must be in read-only format.
\item 	Changes must be approved by the Speaker and Deputy-Speaker.
 
\end{longenum}

\subsubsection{ 3.6.3. Documents}
\begin{longenum}[ label*=\thesubsubsection.\arabic*., align=left]
\item   	Council must approve Bylaws and Constitutions Documents. 
\item 	Council must approve the Equity and Issues Policy Document.
\item 	Council must approve the Orientation Committee Policy document. The last paragraph must use letters (i.e. C and D), instead of numbers, for clarity.
\item 	Personal e-mails will be excluded from council packages and documents, unless given written consent. 

\end{longenum}

\subsection{3.7  Annual General Meeting}
\begin{enumerate}
\item Committee chairs must present updates/information at the Anual General Meeting (AGM). Failure to do so will result in a reduction in honourarium, and chairs are required to present at the next SOGS meeting.
\end{enumerate}

\subsection{3.8 Canadian Federation of Students (CFS)}
\begin{enumerate}
\item  	The society remains neutral to the CFS referendum.
\item 	The society joins other student unions in signing the CFS ‘ Release the Text’ ultimatum. 
\end{enumerate}

\subsection{3.9 London Transit Commission (LTC)}
\begin{enumerate}
\item  	 	The President and VP Student Services are responsible for all negotiations with the LTC. 
\end{enumerate}

\subsection{3.10 Health Plan}
\begin{enumerate}
\item  	Members have the right to opt-out, without undue barriers.
\item	The society will align the fall out-in/out deadlines with the USC’s.
\item	The societies health plan shall be retroactive, and be extended up to 10 days after the deadline, with the same fees.
\item	Students on medical, maternity, and compassionate leave have the option to opt-in to the Health Plan.
\item	Students can opt-in due to life-changing circumstances. 
\item	Health plan cost reductions can be granted to students requiring only 6 months of coverage, based on enrollment.
\item	A minimum of 3 Health Care Plan bids must be proposed and decided on by Council.
\end{enumerate}


\subsection{3.11 General Operations Policies}
\begin{enumerate}
\item  The society opposes private, for-profit Universities in Ontario.
\item  The society opposes limiting students work to 10 hours per week.
\end{enumerate}

\section{Committee Policies}
\subsection{4.1 Policy Committee (PC)}
\begin{enumerate}
\item   The PC was directed by Council to create a policy document on XX. 
\item 	The PC Chair will request standing committees to submit policies during the policy manual review. (ID: 1524)
\item 	The PC will take over all policy-related tasks not covered by the BCC. 
\item 	The PC will create operations policies for all other committees.
\end{enumerate}

\subsection{Health Plan Committee}
\begin{enumerate}
\item A Student Health Services representative will be granted ex-officio membership. 
\end{enumerate}

\section{Additional Motions without a home}
\begin{enumerate}
\item	The executive makes all website related decisions until an official policy exists. 
\end{enumerate}
