\section{Bylaws and Constitution Committee}

\subsection{Ratified}
April 19, 2018

\subsection{Mandate}
As per Bylaw 2.9.4.1, the Bylaws and Constitution Committee (BCC) is a standing committee of the Society of Graduate Students. As per the Society's Bylaws 2.9.5.3, the committee is responsible for devising new and reviewing old bylaws to reflect the current needs and wishes of the Society, and to allow the specific elaboration, interpretation, and application of the Society's Constitution. The Bylaws and Constitution Committee works in conjunction with the Policy Committee to ensure consistency and cohesion within the Society's unified governing documents.

\subsection{Article 1: Membership}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The Bylaws and Constitution Committee shall include:
\begin{enumerate}[label*=\arabic*., align=left]	
\item An official liaison, who shall be the Speaker, and who shall be a non-voting member of the committee in compliance with bylaw 2.9.3.2 and 2.9.3.3.
\item A Chairperson, who shall be chosen from amongst the committee members, and whose duties and responsibilities are outlined in bylaws 2.9.2. The Chairperson of the BCC shall also:
\begin{enumerate}
\item serve as Deputy Speaker (see bylaws 2.2.8)
\item serve, ex-officio, as Deputy Chief Returning Officer (see bylaws 2.2.4)
\item be a voting, ex-officio member of the Policy Committee.
\end{enumerate}
\item A Deputy Chairperson, in compliance with bylaw 2.9.5.3.6
\item The Chairperson of the Policy Committee, who shall be a voting ex–officio member.
\end{enumerate}
\item Other voting members of the BCC shall be full and associate members of the Society nominated to the committee by the Society's Council.
\item Membership, including the Chairperson and Deputy Chairperson, but excluding the ex-officio members, shall be capped at 20 (as per the request of the Society Council).
\end{enumerate}

\subsection{Article 2: Committee Management}
\begin{enumerate}[label*=\arabic*., align=left]	
\item Committee Meetings
\begin{enumerate}[label*=\arabic*., align=left]	
\item The BCC shall endeavour to meet on a monthly basis in order to ensure that the governing documents of the Society are regularly reviewed.
\item The BCC shall endeavour to schedule meetings in the week prior to the deadline for submissions to the Society's Council Package, in order to ensure that any motions pertaining to the governing documents of the Society are afforded the opportunity for review.
\item The BCC will act in accordance with Bylaw 2.9.3.9 to manage and maintain attendance.
\end{enumerate}

\item Voting Policies
\begin{enumerate}[label*=\arabic*., align=left]	
\item The BCC  will act in accordance with Bylaw 2.9.3.5 and 2.9.3.7 to determine quorum, and with Bylaw 2.9.4.2.1 to count votes.
\item Committee members, including Ex-officio, can hold up to three (3) proxies.
\item Aside from the scheduled meetings, online voting is also permitted on the condition that it is for fixing spelling, grammar, or wording errors. Voting members shall be given one (1) business day to respond to an online vote, at which point the vote shall be considered concluded. If any voting member objects to business being voted online, the voting shall be prohibited for that particular piece of business.
\end{enumerate}

\end{enumerate}

\subsection{Article 3: Governing Documents}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The BCC shall conduct a yearly review of the Society's Constitution and Bylaws to ensure consistency and cohesion within the Society's unified governing documents.
\item Constitution of the Society: Amendments to the Society's Constitution will follow the directives outlined in section 1.8.1 (Amendments to the Constitution).
\item Bylaws of the Society: Amendments to the Bylaws shall be dealt with in the same manner as amendments to the Constitution, except that they need not be approved by a General Meeting before coming into effect.
\item Operations and Policies of the Society: The BCC shall support the Policy Committee to maintain the Policy documents of the Society, as well as in drafting motion-based policies.
\end{enumerate}

\subsection{Article 4: Dissemination of Amendments}
\begin{enumerate}[label*=\arabic*., align=left]	
\item In accordance with the Constitution of the Society (1.8.3), amendments to the Constitution that have been ratified at a General Meeting or updates to the Bylaws or Policy Documents that have been ratified at Council shall be both published in the complete form of their respective document and announced to the membership within two (2) months of ratification.
\end{enumerate}