\section{GradCast Editorial Board}

\textbf{Mandate:} A Standing Committee (Bylaws 2.9.5.9) of the Society, the mandate of the GradCast Editorial Board is to highlight graduate students at the University of Western Ontario in a media context.

This policy is a mandatory list of initiatives that the GradCast Editorial Board must undertake in any given year. Here, the aim is to ensure a successful organization of student guests on the CHRW radio talk show and podcast formally known as GradCast.

\subsection{Article 1: Membership} 
\begin{enumerate} [label*=\arabic*., align=left]	
\item The Gradcast Editorial Board shall: 
\begin{enumerate}
\item have an Official Liaison, who shall be the Vice-President Academic, in compliance with the Society's bylaw 2.9.3.2 
\item have a Chairperson who shall:
\begin{enumerate}
\item be known as the Managing Editor who shall act in compliance with the Duties of the Chairperson as outlined in the Society's bylaws 2.9.2, and 
\item send out reminders to guest one month and one week prior to recording.
\end{enumerate}

\item  aim to include at least one member of the Society from each of the four divisions: Arts, Biosciences, Physical Sciences, and Social Sciences.
\end{enumerate}

\item Membership shall be limited to fifteen (15) members, including the Official Liaison and Chair.
\item Positions within the GradCast Editorial Board shall include:
\begin{enumerate}
\item A Promotions, Advertisement and Social Media Manager, who will be responsible
for actively maintaining a social media presence for GradCast
\item A Recruitment Officer, who will be responsible for actively recruiting students to the
show.
\item A Recording Secretary, who will be responsible for keeping and disseminating
committee meeting minutes.
\end{enumerate}
\item The Chair position will be reviewed and voted upon each April, with a transition after a second vote in September
\item Meetings are comprised of members of the Editorial Board and, by invitation by the Chair or request by a Committee member only, non-committee members may attend as non-voting guests.
\item Voting can be conducted by email, telephone, or text.
\item Respectful Abandonment: In the case a member of the editorial board does not attend meetings
or shows for two (2) consecutive weeks without explanation, an email will be sent to the member asking for a follow-up. If they do not respond with a reason within another two (2) weeks, the the chair may, at the discretion of the other members of the committee, remove the member from the committee.
\end{enumerate}


\subsection{GradCast Radio Show and Podcast}
\begin{enumerate} [label*=\arabic*., align=left]	
\item The GradCast Editorial Board shall:
\begin{enumerate}
\item Advertise for submissions in compliance with bylaw 2.9.5.9.3
\item Provide instructions for contributors––episode producer(s), guest(s), and host(s)––in
compliance with bylaw 2.9.5.9
\item Maintain a schedule of guests on a shared online document
\item Once a term, or as needed, host a recruitment event
\item Endeavor to promote GradCast episodes through all available channels
\item Ensure the quality of each GradCast episode
\item Display the official logo for GradCast on all websites and social media platforms
\item Designate at least one (1) member to complete all mandatory CHRW training
\item Designate a producer for each episode, who shall:
\begin{enumerate}
\item Be present at the recording session and set up recording equipment
\item Instruct the host(s) and guest(s) on proper recording etiquette
\item Within one week from recording, edit the podcast and include the GradCast jingle to the beginning and end of the episode
\item Upload each episode to Podbean with title, summary, and photo
\item Upload each episode to CHRW’s shared folder for the station to air
\end{enumerate}
\end{enumerate}
\item The GradCast radio show and podcast shall aim to:
\begin{enumerate}
\item highlight and showcase student research
\item appeal to various audiences including other graduate students, the Western University
campus, the general community
\item maintain episodes at approximately 30 minutes in length on a weekly basis or as
designated by CHRW
\item maintain at least one (1) pre-recorded podcast episode to maintain a weekly schedule
as needed should circumstance arise.
\item be comprised of at least one (1) student guest per episode.
\end{enumerate}
\item The GradCast website shall:
\begin{enumerate}
\item be updated on a regular basis to keep up to date with guest appearances,
\item bear the official GradCast logo,
\item aim to include a picture of guests and hosts, along with a brief caption summarizing the topic
of the episode.
\end{enumerate}
\end{enumerate}