\section{Health Plan Committee}

\subsection{Ratified}
March 13, 2018

\subsection{Mandate}
As per Bylaw 2.9.4.1, the Health Plan Committee (HPC) is a standing committee of the Society of Graduate Students (SOGS). As per the Society's Bylaws (2.9.5.13), the HPC must: review and recommend changes to the Society's health plan and health plan policies, review health plan bids and contracts, provide shortlists to Council, and work in consultation with the Vice-President Services to help run and promote mental health and wellness initiatives (e.g. Graduate Wellness Week) for graduate students.

\subsection{Article 1: Membership}
\begin{enumerate} [label*=\arabic*., align=left]	
\item Chairperson: A person who is selected by the committee.
\item Deputy Chairperson (if applicable): A person who is selected by the committee.
\item Other voting members: 10 other members that are elected to the Committee by the HPC and then ratified
by SOGS Council as per Bylaw 2.9.2.3. All members are bound by Bylaw 2.9.3.9.
\item Non-voting members: Vice-President Student Services (Official Liaison, Ex-officio), President (Ex-officio),
and SOGS Accounting Manager (Ex-officio), and Communications Administrator (Ex-officio). 
\index{Vice--President Student Services!Ex-Officio!Health Plan Committee}
\index{Vice--President Student Services!Health Plan}
\item The above shall remain in accordance with Bylaws 1.6.4.2, 2.9.3.7, and 2.9.3.8.
\item Membership Limit: As per Bylaw 2.9.4.2.2, membership limit for the HPC is set at 12 voting members, which includes the Chairperson, Deputy Chairperson, and 10 other voting members.
\end{enumerate}

\subsection{Article 2: Committee Management}
\begin{enumerate} [label*=\arabic*., align=left]	
\item Committee Archiving: The Communications Administrator will submit HPC minutes to the SOGS Public
Dropbox and the Vice-President Student Services will store HPC minutes and relevant documents in the SOGS VP Student Services Google Drive folder under "Health Plan Committee."
\item General Policies
\begin{enumerate} [label*=\arabic*., align=left]	
\item The HPC will act in accordance with Bylaw 2.9.3.5 and 2.9.3.7 to determine quorum; Bylaw 2.9.3.9 to manage and maintain attendance; and Bylaw 2.9.4.2.1 to count votes.
\item Committee members, including Ex-officio, can hold up to three proxies.
\end{enumerate}
\item HPC Policies
\begin{enumerate} [label*=\arabic*., align=left]	
\item For general voting via email, members will "Reply All" to the HPC listserv.
\item For private votes (e.g. elections), votes will be sent to the Chair and Deputy Chair.
\item oting members shall be given one week to respond (unless in urgent circumstances) to
an email vote at which point the vote shall be considered concluded.
\item If any voting member objects to business being voted on by email, email voting shall be
prohibited for that particular piece of business.
\item Email voting shall be allowed under the following circumstances:
\begin{enumerate} [align=left]	
\item For correcting wording or mistakes in documents already approved by the committee. 
\item New motions that are time sensitive.
\item Opt-out requests.
\end{enumerate}
\end{enumerate}
\item The Roles of HPC Ex-Officio and Voting Members:
\begin{enumerate} [label*=\arabic*., align=left]	
\item President
\begin{enumerate}
\item Shall be a non-voting member, Ex-officio, of the HPC.
\item Shall, with the Vice-President Student Services, be a signing authority on the Health Plan Contract.
\end{enumerate}

\item Vice-President Student Services
\begin{enumerate}
\item Shall be a non-voting member, Official Liaison and Ex-officio, of the HPC.
\item Shall, with the President, be a signing authority on the Health Plan Contract.
\item Shall be bound by Bylaws 2.1.10 and 2.1.2.4.
\item Shall oversee the administration of the Health Plan in consultation with the President, the Health Plan Committee (HPC), and the Plan Broker.
\item Shall be the primary contact person for Plan Bidders in a tendering year.
\item Shall work with the Accounting Manager, Administrative Assistant, and Communications Administrator to maintain website Health Plan information.
\end{enumerate}
\item HPC Chair
\begin{enumerate}
\item See SOGS Bylaws 2.9.2 for the general duties of a SOGS Committee Chair.
\item Shall act in accordance with Bylaw 2.9.5.13 ("Health Plan Committee").
\item Shall call monthly meetings (unless otherwise agreed upon by the committee).
\begin{enumerate}
\item Proposed dates should be sent out at least two weeks before the required meeting.
\item Members should be given at least one week's notice prior to the meeting date.
\item Members should be provided with multiple options for meeting times (e.g.
evenings).
\end{enumerate}

\item Shall assign the Deputy Chair or another Committee member to take meeting minutes in
accordance with Society bylaws, "Minutes Writing Policy."
\item Shall distribute Committee minutes via the HPC listserv as per Bylaw 2.9.2.8.1.
\item Shall submit HPC updates, AGM Reports, Final Reports, Surveys, and Health Plan Bid
Shortlists, as approved by the HPC, for consideration at SOGS Council as per Bylaws
2.15.3.7 and 2.15.3.7.1.
\item Shall liaise with the Vice-President Student Services, President, and Broker for any and all
work related to the Health Plan (contracts, surveys, town halls, etc.).
\item Shall update the HPC Chair Transitional Manual at the end of their term, and submit it for
review to the Vice-President Student Services before stepping down and/or graduating.
\end{enumerate}
\item Committee members
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.3.9.
\item Shall regularly attend committee meetings to the best of their ability.
\item Shall participate in the committee's projects to the best of their ability.
\item Shall vote to approve minutes and motions, and in elections (e.g. HPC Chair).
\end{enumerate}
\end{enumerate}
\item Maintaining Membership
\begin{enumerate} [label*=\arabic*., align=left]	
\item Communications Administrator: Oversees the HPC listserv and committee member list.
\item VP Student Services: Ratifies Committee members at Council and informs the Chair about new members.
\item Chair: Enforces SOGS Committee policies as per Bylaws 2.9.3. and 2.9.4, and informs the Vice-President Student Services when new members are added, and when members resign.
\item In the situation where the Chair position becomes vacant:
\begin{enumerate}
\item The HPC Deputy Chairperson serves as chair pro tempore until the next HPC meeting where an election to replace the Chairperson must take place.
\item If for whatever reason, the Chair is unavailable temporarily, the Deputy Chair will fulfil the roles and responsibilities as delegated by the Chair until the Chair returns.
\item The only circumstances in which the HPC may be temporarily chaired by the Official Liaison is for the purpose of conducting an election to select an existing member of the HPC to the position of Chair even if for a temporary, non-permanent and short-term basis, or until the Chair position is filled.
\end{enumerate}
\item Committee Members
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.3.9.
\item Shall inform the Chair and Vice-President Student Services, and copy the
Communi- cations Administrator, if they wish to resign from the committee.
\end{enumerate}
\end{enumerate}
\end{enumerate}

\subsection{Article 3: Health Plan Opt-Ins / Opt-Outs}
\begin{enumerate}[label*=\arabic*., align=left]	
\item Opt-In/Opt-Out Deadlines: Health Plan opt-ins/opt-outs shall be permitted until the last day of class of the first semester in which a graduate student becomes enrolled at Western University.
\item Opt-In/Opt-Out Appeals: Health Plan opt-ins/opt-outs shall be permitted after the deadline under the following circumstances/conditions:
\begin{enumerate}[label*=\arabic*., align=left]
\item Proof of completion of non-full-time student status, such as completion of studies.
\item Proof of alternate health plan coverage (not including UHIP, OHIP, or provincial equivalent).
\item Under the discretion of the HPC.
\end{enumerate}
\end{enumerate} 

\subsection{Article 4: Health and Dental Plan}
\begin{enumerate} [label*=\arabic*., align=left]	
\item The HPC Official Liaison and the President are responsible for ensuring these policies are enforced.
\item In a non-tendering year, the HPC will devote its time to research and data collection in order to continue to improve the Health and Dental Plan.
\item General Procurement Policy
\begin{enumerate}[label*=\arabic*., align=left]	
\item The HPC shall act in accordance with Society bylaws on "Procurement."
\item The HPC shall specifically act in accordance with Society bylaws on "Request for Proposal."
\item The HPC shall sign a non-disclosure agreement for the duration of the process until the
contract is executed and signed.
\end{enumerate}
\item Health Plan Procurement Policy
\begin{enumerate} [label*=\arabic*., align=left]	
\item The HPC shall act in accordance with Society bylaws on the "Health Plan."
\item The HPC shall act in accordance with Society bylaws on "Conflict of Interest."
\end{enumerate}
\item Health \& Dental Plan Timeline
%\begin{enumerate}[label*=\arabic*., align=left]	
%\item Prepare to Tender. October and November. See archived documents/data.
%\item Circulate RFP. First week in December. See Procurement Policy.
%\item Receive Bidder Questions. Second week in January. See archived documents.
%\item Respond to Questions. Third week in January. See archived documents.
%\item Proposal/Bid Deadline. Second week of February. See Procurement Policy.
%\item Evaluate Proposals/Bids. Third week of February. See archived documents.
%\item Bidder Interviews. Fourth week of February. See archived documents.
%\item Submit Shortlist to Council. Last Thursday of March (Council). See Bylaw 3.5.17.2.
%\item Final Report to Council. Last Thursday of May (Council). See archived documents.
%\end{enumerate}

\begin{tabular}{|l|l|l|}
\hline 
\multicolumn{3}{|l|}{\textbf{Milestone}} \\ 
\hline 
Prepare to Tender & October and November & See archived data \\ 
\hline 
Circulate RFP & First week in December & See Procurement Policy \\ 
\hline 
Receive Bidder Questions & Second week in January & See archived documents \\ 
\hline 
Respond to Questions & Third week in January & See archived documents \\ 
\hline 
Proposal/Bid Deadline & Second week of February & See Procurement Policy \\ 
\hline 
Evaluate Proposals/Bids & Third week of February & See archived documents \\ 
\hline 
Bidder Interviews & Fourth week of February & See archived documents \\ 
\hline 
Submit Shortlist to Council & Last Thursday of  March (Council) & See Bylaw 3.5.17.2 \\ 
\hline 
Final Report to Council & Last Thursday of \newline May (Council) & See archived documents \\ 
\hline 
\end{tabular} 

\item Health \& Dental Plan Award Process

\begin{enumerate}[label*=\arabic*., align=left]	
\item Involves vetting of submitted bids in addition to a 45-minute interview of each bidder.
\item The HPC will construct its shortlist based on the following criteria:
\begin{enumerate}
\item Price -- 30\% (e.g. Lowest premium)
\item Specific Compliance -- 25\% (e.g. financial responsible model(s) for SOGS)
\item Quality -- 20\% (overall financial and service value to graduate students; coverage)
\item Terms and Conditions -- 10\% (e.g. blackout period, dispensing fee, family opt-ins
fees)
\item Experience -- 10\% (e.g. working directly with other Graduate Societies)
\item Other -- 5\% (Discounts, Network, Bursaries, Sponsorship, Promotional,
Communication, Administrative, Data Analytics, Flex Plans, Mental health/Wellness initiatives, etc.).
\end{enumerate}
\end{enumerate}
\end{enumerate}

\subsection{Article 5: Graduate Student Mental Health and Wellness}
\begin{enumerate}[label*=\arabic*., align=left]	
\item Work in partnership with the Vice-President Student Services, Graduate Peer Support (GPS), and the Health
\& Dental Plan Broker in preparation for Graduate Wellness Week (October) and its initiatives.
\item During non-tendering years, the HPC will work in partnership with other SOGS committees to advocate on
behalf of graduate students in terms of their mental health and wellness.
\end{enumerate}