\section{International Graduate Student Issues Committee}
\subsection{Ratified}
April 18, 2018
\subsection{Mandate}
As per Bylaw 2.9.4.1, the International Graduate Student Issues Committee (IGSIC) is a standing committee of the Society of Graduate Students (SOGS). As per the Society's Bylaws (2.9.5.7), IGSIC must: assess the particular needs of international graduate students, advocate on behalf of them, enhance their overall student life and community at Western University, work in consultation with the Vice-President Student Services to run advocacy, social, and wellness initiatives for international graduate students, and sometimes work in partnership with the International and Exchange Student Centre (IESC), the School of Graduate and Postdoctoral Studies (SGPS), and other relevant SOGS committees.
\subsection{Article 1: Membership}
\begin{enumerate}  [label*=\arabic*., align=left]
\item Chairperson: A person who is selected by the committee.
\item Other voting members: 19 other members that are elected to the Committee by IGSIC and
then ratified by SOGS Council as per Bylaw 2.9.2.3. All members are bound by Bylaw 2.9.3.9.
\item Non-voting members: Vice-President Student Services (Official Liaison, Ex-officio),
President (Ex-officio), and Communications Administrator (Ex-officio).
\item The above shall remain in accordance with Bylaws 1.6.4.2, 2.9.3.7, and 2.9.3.8.
\item Membership Limit: As per Bylaw 2.9.4.2.2, membership limit for IGSIC is set at 20 voting
members, which includes the Chairperson and 19 other voting members.
\end{enumerate}

\subsection{Article 2: Committee Management}
\begin{enumerate}[label*=\arabic*., align=left]
\item Committee Archiving: The Communications Administrator will submit IGSIC minutes to the
SOGS Public Dropbox and the Vice-President Student Services will store IGSIC minutes and relevant documents in the SOGS Vice-President Student Services Google Drive folder under "International Graduate Student Issues Committee (IGSIC)."
\item Voting Policies
\begin{enumerate}[label*=\arabic*., align=left]
\item IGSIC will act in accordance with Bylaw 2.9.3.5 and 2.9.3.7 to determine quorum; Bylaw 2.9.3.9 to manage and maintain attendance; and Bylaw 2.9.4.2.1 to count votes.
\item For general voting via email, members will "Reply All" to the IGSIC listserv.
\item For private votes (e.g. elections), votes will be sent to the Chair, or the Vice-President Student Services if the position is vacant.
\end{enumerate}
\item IGSIC Policies
\begin{enumerate}[label*=\arabic*., align=left]
\item For volunteering at IGSIC Orientation, events, and/or trips, IGSIC will:
\begin{enumerate}
\item Subsidize 50\% of the volunteer ticket for off-campus IGSIC trips.
\item Offer a free ticket for Grad Club events involving meals.
\item Staff up to a maximum of two volunteers/per bus for trips, and up to a maximum of 4 volunteers for events (determined by committee vote).
\end{enumerate}
\item IGSIC will attend the IESC's Global Cafe once a term: September (Orientation: third week), Winter (third one in January), and Summer (third one in May).
\item For IGSIC attendance, members are bound to SOGS bylaws, in addition to:
\begin{enumerate}
\item To be considered an active member of IGSIC, you must attend 50\% of the
committee's meetings and/or actively/regularly participate in the
committee's events and initiatives.
\item Should a member continually not assign proxies/not participate in events
and/or initiatives, IGSIC will vote to remove the member after issuing a first
warning.
\end{enumerate}
\end{enumerate}
\item The Roles of IGSIC Ex-Officio and Voting Members
\begin{enumerate}[label*=\arabic*., align=left]
\item The President shall be a non-voting member, ex-officio, of IGSIC.
\item Vice-President Student Services 
\begin{enumerate}
\item shall be a non-voting member, Official Liaison and ex-officio, of IGSIC.
\item Shall be bound by Bylaws 2.1.10 and 2.1.2.4.
\item shall work with the Communications Administrator to maintain website
international student website information and the IGSIC Facebook group.
\end{enumerate}
\item IGSIC Chair (See SOGS Bylaws 2.9.2 for the general duties of a SOGS Committee Chair).
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.5.8 ("International Graduate Student Issues Committee").
\item Shall call monthly meetings (unless otherwise agreed upon by the committee)
\begin{enumerate}[label*=\arabic*., align=left]
\item Proposed dates should be sent out at least two weeks before the required
meeting.
\item Members should be given at least one week's notice prior to the meeting date.
\item Members should be provided with multiple options for meeting times (e.g. evenings).
\end{enumerate}
\item Shall assign a Committee member to take meeting minutes in accordance with
Society bylaws, "Minutes Writing Policy."
\item Shall distribute Committee minutes via the IGSIC listserv as per Bylaw 2.9.2.8.1.
\item Shall submit IGSIC updates, AGM Reports, Final Reports, and Surveys, as
approved by IGSIC, for consideration at SOGS Council as per Bylaws 2.15.3.7 and
2.15.3.7.1.
\item Shall liaise with relevant administrative figures (IESC, SGPS, etc) and other
relevant SOGS Committee (Equity, GSIC, OSC, etc) to enhance international
student life.
\item Shall update the IGSIC Chair Transitional Manual at the end of their term, and
submit it for review to the Vice-President Student Services before stepping down
and/or graduating.
\end{enumerate}
\item Committee Members
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.3.9.
\item Shall regularly attend committee meetings to the best of their ability.
\item Shall participate in the committee's projects to the best of their ability.
\item Shall vote to approve minutes and motions, and in elections (e.g. IGSIC Chair).
\end{enumerate}
\item Maintaining Membership:
\begin{enumerate}
\item Communications Administrator oversees the IGSIC listserv and committee member list.
\item Vice-President Student Services ratifies committee members at Council and informs the IGSIC Chair about new members.
\item In the situation where the Chair position becomes vacant, the Vice-President Student Services will hold the position of Chair until one is elected.
\item Chair enforces SOGS Committee policies as per Bylaws 2.9.3. and 2.9.4, and informs the Vice-President Student Services when new members are added, and when members resign.
\item Committee members shall inform the Chair and Vice-President Student Services, and copy the
Communications Administrator, if they wish to resign from the committee.
\end{enumerate}
\end{enumerate}
\end{enumerate}

\subsection{Article 3: Fall Orientation}
\begin{enumerate}[label*=\arabic*., align=left]
\item IGSIC, in partnership with the Vice-President Student Services, Grad Club Manager, and the
OSC, will organize and run SOGS' IGSIC Orientation, which usually falls on the Tuesday of the third week of September. This will also include partnering with the Grad Life Advisory Group.
\item IGSIC will begin planning and advertising Fall Orientation in April and/or May.
\item For detailed information on IGSIC Orientation, the Chair should refer to the IGSIC Chair
Transitional Manual and IGSIC should refer to the IGSIC Google Drive Archive overseen by
the Vice-President Student Services.
\end{enumerate}

\subsection{Article 4: Socials and Trips}
\begin{enumerate}[label*=\arabic*., align=left]
\item IGSIC shall work in conjunction with the Vice-President Student Services, the International and Exchange Student Centre (IESC), and the School of Graduate and Postdoctoral Studies (SGPS) to host social events and trips for international graduate students whenever possible.
\item IGSIC shall endeavor to offer, organize, and run family-friendly and/or non-alcohol focused events in an effort to better meet the needs/interests/ of international graduate students.
\item IGSIC will host at cost and/or subsidized off-campus trips each month of the summer for international graduate students to the best of their ability.
\item For detailed information on event history, IGSIC should refer to the IGSIC Google Drive Archive overseen by the Vice-President Student Services.
\end{enumerate}

\subsection{Article 5: Advocacy and Partnerships}
\begin{enumerate}[label*=\arabic*., align=left]
\item IGSIC shall endeavor to advocate on behalf of international graduate students through
surveys, petitions, services, and information sharing initiatives and events.
\begin{enumerate}[label*=\arabic*., align=left]
\item IGSIC shall endeavor to conduct an International Graduate Student Satisfaction
Survey every two years to ensure the needs of its constituents are being met.
\item IGSIC will endeavor to advocate on behalf of one-year programs by reaching
out to them and meeting with them on a case-by-case basis.
\item IGSIC shall endeavor to, when possible, partner with SOGS, PSAC 610, and the
Canadian Federation of Students (CFS) to enhance the experiences of international
graduate students.
\item IGSIC shall endeavor to host Immigration Seminars at least twice a year in partnership with external legal experts and/or the IESC.
\item IGSIC shall endeavor to host at least one International graduate student specific Tax Information session in the Winter term in partnership with the IESC.
\end{enumerate}
\item IGSIC shall endeavor to partner with other SOGS Committees whenever possible; for example, Equity Committee, Graduate Student Issues Committee (GSIC), the Orientation and Social Committee (OSC), etc.
\item IGSIC, through the Vice-President Student Services, shall endeavour to partner with graduate student stakeholders (e.g. Graduate Student Life Advisory Group) for IGSIC Orientation and international graduate-specific events.
\end{enumerate}