\section{Orientation and Social Committee}
\index{Committee Policy!Orientation and Social Committee}
\index{Committee!Orientation and Social Committee!Committee Policy}
\subsection{Ratified}
March 5, 2018
\subsection{Mandate}
As per Bylaw 2.9.4.1, the Orientation and Social Committee (OSC) is a standing committee of the Society of Graduate Students (SOGS). As per the Society's Bylaws (2.9.5.8), the OSC must work in consultation with the Vice-President Student Services to: organize and facilitate SOGS Fall, Winter, and Summer Orientations, host social and recreational--including non-alcoholic--events for SOGS' diverse membership, work in conjunction with the Communications Administrator, Grad Club Manager, SGPS, and other SOGS Committees and Graduate Associations when necessary.

\subsection{Article 1: Membership}
\begin{enumerate}[label*=\arabic*., align=left]	
\item Chairperson: A person who is selected by the committee. When there are two Chairs, they will be Co-Chairs and will split the Chair honorarium. 
\item Other voting members: 10 other members that are elected to the Committee by the OSC and then ratified
by SOGS Council as per Bylaw 2.9.2.3. All members are bound by Bylaw 2.9.3.9.
\item Non-voting members: Vice-President Student Services (Official Liaison, Ex-officio), President (Ex-officio),
Communications Administrator (Ex-officio), and any members from other SOGS committees who have been
invited by the Official Liaison.
\item The above shall remain in accordance with Bylaws 1.6.4.2, 2.9.3.7, and 2.9.3.8.
\item Membership Limit: As per Bylaw 2.9.4.2.2, membership limit for the OSC is set at 12 voting members, which includes the Chairperson(s) and 10 other members (11 if there is one Chair).
\end{enumerate}

\subsection{Article 2: Committee Management}
\begin{enumerate}[label*=\arabic*., align=left]	
\item Committee Archiving: The Communications Administrator will submit OSC minutes to the SOGS Public
Dropbox and the Vice-President Student Services will store OSC minutes and relevant documents in the SOGS VP Student Services Google Drive folder under "Orientation and Social Committee."
\item General Voting Policies
\begin{enumerate}[label*=\arabic*., align=left]	
\item The OSC will act in accordance with Bylaws 2.9.3.5 and 2.9.3.7 to determine quorum (50\% + 1), Bylaw 2.9.3.9 to manage and maintain attendance, and Bylaw 2.9.4.2.1 to count votes.
\item Committee members, including Ex-officio, can hold up to three proxies.
\end{enumerate}
\item OSC Voting Policies
\begin{enumerate}[label*=\arabic*., align=left]	
\item For general voting via email, members will "Reply All" to the OSC listserv.
\item For private votes (e.g. elections), votes will be sent to the Chair, or the Vice-President Student Services if the position is vacant.
\item For volunteering at OSC Orientation, events, and/or trips, the OSC will:
\begin{enumerate}
\item Subsidize 50\% of the volunteer ticket for off-campus OSC trips.
\item Offer a free ticket for Grad Club events involving meals.
\item Staff up to a maximum of two volunteers/per bus for trips, and up to a maximum of 5
volunteers for events (determined by committee vote).
\end{enumerate}
\item The OSC will host and oversee Board Game Night at the Grad Club on the third Wednesday of every month and staff at least two volunteers to run the event.
\end{enumerate}
\item The Roles of OSC Ex-Officio and Voting Members
\begin{enumerate}[label*=\arabic*., align=left]	
\item President
\begin{enumerate}
\item shall be a non-voting member, Ex-officio, of the OSC.
\end{enumerate}
\item Vice President Student Services
\begin{enumerate}
\item Shall be a non-voting member, Official Liaison and Ex-officio, of the OSC.
\item Shall be bound by Bylaws 2.1.10 and 2.1.2.4.
\item Shall oversee Chair and committee member training with the Speaker in April and provide OSC uniforms.
\item Shall oversee the execution of Fall Orientation in consultation with the IGSIC Chair and Committee, which includes logistics, sponsorship, and advertising.
\item Shall introduce IGSIC to relevant IESC and SGPS administrative contacts.
\item Shall work with the Accounting Manager to ensure that IGSIC stays within its budget.
\item Shall work with the Communications Administrator to ensure the successful
promotion and visibility of IGSIC Orientation and Events (e.g. SOGS website, "Orientation," SOGS Orientation Handbook, SOGS social media, including IGSIC Facebook group, SOGS newsletter, International Graduate Student Listserv, etc.).
\end{enumerate}
\item Communications Administrator
\begin{enumerate}
\item Purchases supplies for Orientation and events when given one month's notice.
\item Designs graphics, tickets, and advertises (social media, newsletter, website) for all events when given at least two weeks' notice.
\end{enumerate}
\item OSC Chair
\begin{enumerate}
\item See SOGS Bylaws 2.9.2 for the general duties of a SOGS Committee Chair.
\item Shall act in accordance with Bylaw 2.9.5.8 ("Orientation and Social Committee").
\item Shall meet with the Speaker and Vice-President Student Services in April for training.
\item Shall meet with the new members as they join the committee, and introduce them to
the Grad Club Manager.
\item Shall call monthly meetings (unless otherwise agreed upon by the committee).
\begin{enumerate}[label*=\arabic*., align=left]	
\item Proposed dates should be sent out at least two weeks before the required meeting.
\item Members should be given at least one week’s notice prior to the meeting date.
\item Members should be provided with multiple options for meeting times (e.g. evenings).
\end{enumerate}
\item Shall assign the Co-Chair or another Committee member to take meeting minutes in accordance
with Society bylaws, Minutes Writing Policy.
\item Shall distribute Committee minutes via the OSC listserv as per Bylaw 2.9.2.8.1.
\item Shall inform the Communications Administrator, and copy the Vice-President Student Services, at least one month in advance if committee supplies must be purchased and off-campus events arranged.
\item Shall inform the Communications Administrator, and copy the Vice-President Student Services, at least two weeks in advance regarding event information for advertising, social media, and event ticket design/e-ticking purposes.
\item Shall contact the Grad Club Manager, and copy the Vice-President Student Services, to book Grad Club Events.
\item Shall oversee assigning volunteers for Orientation and/or social events.
\item Shall submit AGM Reports and Final Reports, as approved by the OSC, to SOGS Council as per Bylaws 2.15.3.7 and 2.15.3.7.1.
\item Shall liaise with the Grad Club Committee, Graduate Peer Support (GPS) and other SOGS Committees in regards to Fall, Winter, and Summer Orientations and Graduate Wellness Week.
\item Shall update the OSC Chair Transitional Manual at the end of their term, and submit it for review to the Vice-President Student Services, before stepping down and/or graduating.
\end{enumerate}
\item Committee members
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.3.9.
\item Shall receive training/documents/uniforms from the Speaker and Vice-President Student Services in April.
\item Shall regularly attend committee meetings to the best of their ability.
\item Shall participate in the committee's events and projects to the best of their ability through
logistical planning and/or staffing an event.
\item Shall endeavour to wear OSC uniform at OSC Orientation and events. d. Shall vote to approve minutes and motions, and in elections (e.g. OSC Chair).
\end{enumerate}

\end{enumerate}
\item Maintaining Membership
\begin{enumerate}[label*=\arabic*., align=left]	
\item Communications Administrator: Oversees the OSC listserv and committee member list.
\item Vice-President Student Services: Ratifies Committee members at Council and informs the
OSC Chair(s) about new members.
\begin{enumerate}
\item In the situation where the Chair position becomes vacant, the Vice-President Student
Services (Official Liaison) will hold the position of Chair until one is elected.
\end{enumerate}
\item Chair(s): Enforces SOGS Committee policies as per Bylaws 2.9.3-2.9.4 and informs the Vice-
President Student Services when new members are added, and when members resign.
\item Committee members:
\begin{enumerate}
\item Shall act in accordance with Bylaw 2.9.3.9.
\item Shall inform the Chair and Vice-President Student Services, and copy the
Communi- cations Administrator, if they wish to resign from the committee.
\end{enumerate}
\end{enumerate}
\end{enumerate}

\subsection{Article 3: Fall, Winter, and Summer Orientations}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The OSC, in partnership with the Vice-President Student Services, Grad Club Manager, and other SOGS
committees, will organize and run SOGS Orientation, which usually runs during the third week of September, the third week of January, and the third week of May. This will also include partnering with SGPS for Graduate Fall, Winter, and Summer Orientations, which usually occur during the first weeks of September, January, and May.
\item The OSC will begin planning Fall Orientation in April, Winter Orientation in November, and Summer Orientation in March.
\item For detailed information on Orientation, the Chair(s) should refer to the OSC Chair Transitional Manual and the OSC should refer to the OSC Google Drive Archive overseen by the Vice-President Student Services.
\end{enumerate}

\subsection{Article 4: Socials and Recreation}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The OSC shall work in conjunction with the Grad Club Manager and the Grad Club Committee to host
events at the Grad Club whenever possible.
\item The OSC shall endeavor to offer, organize, and run social and/or recreational and/or non-alcohol focused
events in an effort to better meet the needs/interests/demographics of graduate students.
\item The OSC will oversee Board Game Night (every third Wednesday of the month) at the Grad Club.
\item For detailed information on event history, the OSC should refer to the OSC Google Drive Archive overseen
by the Vice-President Student Services.
\end{enumerate}

\subsection{Article 5: Graduate Student Life: Partnership and Outreach}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The OSC shall endeavor to partner with other SOGS Committees whenever possible; for example, Grad Club
Committee (Stitch \& Bitch), Graduate Peer Support (Graduate Wellness Week), International Graduate
Student Issues Committee (family friendly, non-alcohol events), etc.
\item The OSC shall endeavor to reach out to representatives from Graduate Associations/Departments regarding
Orientation and social events throughout the year for the purpose of feedback and enhancement of
SOGS/graduate-specific events.
\item The OSC, through the Vice-President Student Services, shall endeavour to partner with graduate student
stakeholders (e.g. Graduate Student Life Advisory Group) for Orientation and graduate-specific events.
\end{enumerate}
