\section{Policy Committtee}
\subsection{Ratified}
15 March 2018
\subsection{Mandate}
A Standing Committee of the Society (Bylaw 2.9.5.11), the Policy Committee is mandated with maintaining the Policy Manual and motions database, as well as drafting motion- based policies. The Policy Committee works in conjunction with the Bylaws and Constitution Committee to ensure consistency and cohesion within the Society's unified governing documents.
\subsection{Membership}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The Policy Committee shall have:
\begin{enumerate}[label*=\arabic*., align=left]	
\item An Official Liaison, who shall be the President, in compliance with Bylaws 2.9.3.2 and
2.9.3.3.
\item A Chairperson, who shall:
\begin{enumerate}
\item be chosen from amongst the committee members,
\item act in compliance with the Duties of the Chairperson (Bylaw 2.9.2), and
\item fulfill the duties in Bylaw 2.9.5.11.1.
\end{enumerate}
\item The Deputy Speaker shall be a voting ex-officio member (Bylaw 2.9.5.11.2).
\item The Ombudsperson shall be a non-voting ex-officio member (Bylaw 2.2.7.5).
\item Membership shall be limited to fifteen (15) members, including the Official Liaison,
Chairperson, and ex-officio members.
\end{enumerate}

\subsection{Article 2: Committee Management}
\begin{enumerate}[label*=\arabic*., align=left]	
\item The Policy Committee shall act in accordance with Bylaw 2.9.3.9 to manage and maintain
attendance and with Bylaws 2.9.3.5 and 2.9.3.7 to determine quorum.
\item Voting Policies
\begin{enumerate}[label*=\arabic*., align=left]	
\item Aside from the scheduled meetings, online voting is also permitted on the condition that it is for correcting spelling, grammar, or wording mistakes on motions. Voting members shall be given one (1) business day to respond to an online vote, at which point the vote shall be considered concluded. If any voting member objects to business being voted on online, the voting shall be prohibited for that particular piece of business.
\item Committee members (including the Official Liaison, Chairperson, and ex-officio members) can hold up to two (2) proxies.
\end{enumerate}
\end{enumerate}
\end{enumerate}
