\section{Other Policies}
\begin{longenum}[label*=\thesection.\arabic*., align=left]
\item Position Statements
\begin{longenum}[label*=\arabic*., align=left]
\item Membership
\begin{longenum}[label*=\arabic*., align=left]
\item The Society supports the right to a leave of absence for its members. Whereas the Society recognizes that departments may have different expectations of graduate students, the terms for leaves of absence are determined by each department separately.
\item The Society supports international students at Western and advocates for fair and regulated international student fees. The Society will lobby and pressure the University to freeze international student tuition fees.
\item The Society supports the right to appeal. In any academic appeal, a graduate student may use the support of students, faculty members, and/or family members.
\item The Society endorses the view that graduate student funding should guarantee cost of living, indexed to CPI
\item The Society encourages the presence of all Executive and Non-Executive Officers at the Western Research Forum (WRF).
\item The Society opposes reductions in vacation time for current employees.
\item Members of the Society shall act in accordance with the Personnel Manual.
\item The Society supports access to the SOGS Health Plan for all Grad Club staff/employees.
\item The Society supports Workers Compensation coverage for its employees. The Society will make budgetary adjustments to match increasing cost of coverage.
\item The Society is committed to paying its Executive wages commensurate with the wages paid to TAs.
\item Whereas the Society supports regulation of student fees and opposes tuition fee increases, the Society supports setting student fees based on 12-month periods.
\item The Society supports the use of external legal professional to resolve contracts issues and impeachment issues. Where the Society is involved in a contract, all matters shall be handled under the advisement of a legal professional.
\item The Society is committed to equality across all departments, and council seats will be determined by SGPS-generated codes. Departments from separate codes are free to form one graduate association, provided they provide the request in writing.
\item The Society remains neutral to CFS referenda.
\item The Society opposes private, for-profit Universities in Ontario.
\item The Society is committed to promoting environmental sustainability.
\item The Society opposes limitations to how many hours students can work per week.
\item The Society condemns all forms of violence against women.
\item The Society is a non-discriminatory organization committed to free speech and equality.
\item The Society supports and embodies the principle of establishment of an independent body to be responsible for student services that is fully accountable to students.
\item The Society is committed to the principle that the Senate Subcommittee of SCAPA should have both graduate and undergraduate student representation.
\item The Society is committed to having graduate students be able to access and use the services of students, faculty members, and or family members in any academic appeal at Western.
\item The Society is principle wishes to have Research Assistants unionized.
\item The Society has no official position regarding the Canadian Federation of Students (CFS) Referendum.
\item The Society supports the creation of open and transparent communication from all organizations it is a member of.
\item The Society will have an official social media presence where appropriate.
\end{longenum}
\item Tuition, Student Fees and Graduate Student Funding
\begin{longenum}[label*=\arabic*., align=left]
\item The Society is committed to guaranteed funding for graduate students.
\item The Society supports regulation and reduction of tuition fees.
\item The Society believes that university education should be affordable for all university students.
\item The Society believes that there should be no distinction in funding and tuition between domestic and international students
\item The Society is committed to the welfare of all graduate students especially in the area of guaranteed funding that is indexed to the CPI to ensure that guaranteed funding reasonably compares with cost of living.
\item The Society supports different ancillary fees for graduate students commensurate with the level of service provided and consistent with the overall level of budget reductions and academic units at Western.
\item The Society supports reducing graduate students' ancillary fees, including fees paid to support Intercollegiate Athletics.
\end{longenum}
\item Deregulation
\begin{longenum}[label*=\arabic*., align=left]
\item The Society in principle opposes deregulation or elimination of government-established limits on tuition fee increases and policies of "tuition fee setting flexibility."
\end{longenum}
\item Other Tuition and Student Fee Issues
\begin{longenum}[label*=\arabic*., align=left]
\item The Society rejects in principle the student support fee (scholarship bursary fee).
\item The Society in principle feels that the Society must be consulted about any increases in the student services fee. The Society further reserves the right to oppose, make recommendations or voice opinions on student services fee increases. The Society feels that such fee increases are not in students' best interests. 
\item The Society is in support of the tuition fees of all graduate students being set upon a 12-month period.
\item The Society supports in principle differential ancillary fees for graduate students consistent with the level of service provided and consistent with the overall level of budget reductions and academic units at Western.
\item The Society is opposed to the use of transcript fees as a source of revenue generation for university operations. 
\end{longenum}
\item Miscellaneous Issues
\begin{longenum}[label*=\arabic*., align=left]
\item The Society is committed to having the British spelling of "Honours" on university degrees.
\item The Society endorses the 1997 change in allocation of research grants.
\end{longenum}
\item Operations Policies
\begin{longenum}[label*=\arabic*., align=left]
\item The Society non-executive officers shall be allocated space in the Society office for the purpose of performing their Society related duties.
\item Special status students, such as exchange and International graduate students, who wish to gain access to services offered by the Society may do so if 1) they pay the society's and the Canadian Federation of Students (CFS) membership fees, and 2) pay all fees normally associated with the Society program(s) they wish to access.
\end{longenum}
\item Finance and Banking
\begin{longenum}[label*=\arabic*., align=left]
\item The Society's Fiscal year ends April 30.
The Society will contribute top-up funds to the Joint Fund for Graduate Scholarship and Research program to match the top-up funding provided by the Faculty of Graduate Studies and Research Western.
\end{longenum}
\item Duties of the Executive
\begin{longenum}[label*=\arabic*., align=left]
\item The Speaker and Committee Members shall review all final reports and office matters regarding all outgoing executive and non-executive members. Upon approval of reports, paychecks will be issued to the parties involved.
\item The Council authorizes the Executive to make investment decisions on behalf of the Society.
\item The Accounting manager has signing authority on all accounts belonging to the Society.
\item The Society will be included in the Grad Club's commercial property and general liability insurance policy, provided an agreement on the insurance broker and premium is agreed upon by both the Society and Grad Club.
\item Executive members of the Society shall be compensated in proportion to TA wages, as follows: 0.5 TA for Vice-President and 1.0 TA for the President.
\end{longenum}
\item Documentation
\begin{longenum}[label*=\arabic*., align=left]
\item Council must approve Bylaws and Constitution Documents.
\item Council must approve the Equity and Issues Policy Document.
\item Audio recordings will be taken to aid the Communications administrator in verifying minutes, and these recordings will not be available outside the Society's office.
\item Personal emails will be excluded from council packages and documents, unless given written consent.
\end{longenum}
\item London Transit Commission (LTC)
\begin{longenum}[label*=\arabic*., align=left]
\item The President and Vice-President Student Services are responsible for all negotiations with the London Transit Commission (LTC).
\item Students who surrender their original buss pass within three weeks of receiving a replacement buss pass are eligible to receive a refund of the replacement charge less a \$5.00 administration fee.
\end{longenum}
\item Committees
\begin{longenum}[label*=\arabic*., align=left]
\item The Finance Committee: The President of the Society will serve as the Society's representatives during negotiations with the University regarding the ongoing working of the Ontario Students Opportunity Trust Fund (OSOTF) programme.
\item Additional Motions without a home: The Executive makes all website related decisions until an official policy exists.
\item Committee policy: Committee chairs must attend and present updates at the SOGS Annual General Meeting. If the Chair fails to do so, they must present an update at the next SOGS meeting and their honorarium will be reduced.
\item Health Plan: The Society will make available restricted funds from the Health plan to assist those students (or our members) that are denied drug coverage after undergoing a full appeal with health plan provider. In those cases, students (members) will receive reimbursement at the same rate as the current health plan provider after review by the Health Plan Committee.
\end{longenum}
\item Constitution and Bylaws
\begin{longenum}[label*=\arabic*., align=left]
\item The constitution and bylaws of the Society must be in read-only format.
\end{longenum}
\end{longenum}
\end{longenum}